---
title: "Colocalization"
author: "Brian M. Schilder"
date: "Most recent update:<br> `r Sys.Date()`"
output:
  BiocStyle::html_document:
vignette: >
    %\VignetteIndexEntry{colocalize}
    %\usepackage[utf8]{inputenc}
    %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=TRUE}
library(catalogueR)
```

## Introduction

Here we aim to robustly test whether the genetic signals underlying two
datasets (e.g. GWAS vs. eQTL in the same locus) are indeed the same,
using a methodology called colocalization.

Specifically, we use
[`coloc`](https://github.com/chr1swallace/coloc), which infers the
probability that each SNP is causal in a given locus in each of the
datasets. It then tests the hypothesis that those signals show
substantially similar association distributions.

## List datasets

We have previously queried *eQTL Catalogue* using several Parkinson's
disease GWAS loci (with `eQTLcatalogue_query()`).

Let's first gather those files saved as TSV files.

```{r import_datasets, eval=FALSE}
gwas.qtl_paths <- catalogueR::eQTLcatalogue_example_queries(fnames = c(
  "BST1__Alasoo_2018.macrophage_IFNg+Salmonella.tsv.gz",
  "BST1__Alasoo_2018.macrophage_naive.tsv.gz",
  "BST1__Alasoo_2018.macrophage_Salmonella.tsv.gz"
))
```

## Run coloc

Since its original release,
[`coloc`](https://github.com/chr1swallace/coloc) has been updated so
that it can now model multiple causal variants within a given dataset
(see the
[paper](https://journals.plos.org/plosgenetics/article?id=10.1371/journal.pgen.1008720)),
whereas previously it could assume only one causal variant. Thus, it may
be able to better estimate the colocalization probability between two
datasets.

```{r run_coloc, eval=FALSE}
coloc_QTLs <- catalogueR::COLOC_run(
  gwas.qtl_paths = gwas.qtl_paths,
  top_snp_only = TRUE,
  split_by_group = FALSE,
  method = "abf")
```

## Plot results

First, let's plot only the results with >80% colocalization probability.
This is a colocalization threshold commonly used in the field.

```{r plot_results, eval=FALSE}
coloc_plot <- catalogueR::COLOC_heatmap(
  coloc_QTLs = coloc_QTLs,
  coloc_thresh = .8)
```

If there are no colocalizations at >=80% probability, you can lower
the threshold to explore the data (though not advisable for real
analyses).

```{r plot_results2, eval=FALSE}
coloc_plot <- catalogueR::COLOC_heatmap(
  coloc_QTLs = coloc_QTLs,
  coloc_thresh = 0)
```

**Too many results:**

In other cases, you may have too many colocalized results to show
in a plot all at once. As a general rule, the more tests you run the
higher your probability threshold should be, since Bayesian methods
do not calculate P-values (and thus cannot use multiple-testing
correction methods like FDR). For example, if you ran many thousands
of coloc tests across many GWAS-eQTL pairs, it is advisable to raise
your PP threshold to 0.90 or even 0.99. This also has the benefit
of narrowing down your (sometimes many!) results so that you may
focus on only the strongest colocalizations.

## Session info

<details>

```{r Session_info}
utils::sessionInfo()
```

</details>
