---
title: "hexSticker"
author: "<h4>Author: <i>Brian M. Schilder</i></h4>" 
date: "<h4>Updated: <i>`r format( Sys.Date(), '%b-%d-%Y')`</i></h4>"
output:
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{hexSticker}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, include=FALSE}
pkg <- read.dcf(here::here("DESCRIPTION"), fields = "Package")[1]
description <- read.dcf(here::here("DESCRIPTION"), fields = "Description")[1]
```

You can make awesome hex stickers for your R packages using:

- [hexSticker](https://github.com/GuangchuangYu/hexSticker)  
- [ggimage](https://github.com/GuangchuangYu/ggimage) 
lets you render images as data points.  
- [ggpattern](https://coolbutuseless.github.io/package/ggpattern/)
lets you fill objects with patterns or images.
- [magick](https://cran.r-project.org/web/packages/magick/vignettes/intro.html) 
modify PNGs.

# `r pkg`
 
```{r setup}
# If you're using R<4.1.1, need this version of rvcheck
# devtools::install_version('rvcheck',version='0.1.8') 
library(hexSticker)
library(dplyr)
library(ggplot2)
library(ggimage)
# library(ggpattern)# remotes::install_github("coolbutuseless/ggpattern")
```

## File path

Create file path.

```{r}
filename <- here::here("inst/hex/hex.png")
dir.create(dirname(filename), showWarnings = FALSE, recursive = TRUE)
```

## Bat logo

Download bat logo from the 
[*echoverseTemplate*](https://github.com/RajLabMSSM/echoverseTemplate/releases/tag/latest)
GitHub Release.

```{r}
tmp <- tempfile()
#### Side view ####
URL <- "https://github.com/RajLabMSSM/echoverseTemplate/releases/download/latest/bat_silhouette.png"   
#### Front view ####
# URL <- "https://github.com/RajLabMSSM/echoverseTemplate/releases/download/latest/bat_silhouette_front.png"   
download.file(URL, tmp)
```

## Background 

Create background with `ggplot2`.

```R
set.seed(1234)
n_bats <- 20
d <- data.frame(x = -rexp(n_bats, rate = 3),
                y = rexp(n_bats, rate = 3)
                ) |>
  dplyr::mutate(image = URL,
                bsize = abs(x*y^2)) |>
  dplyr::arrange(dplyr::desc(x), dplyr::desc(y))
qplot(d$x, d$y, size=d$bsize)
```

Import data points for reproducibility.

```{r}
d <- data.table::fread("https://github.com/RajLabMSSM/echoverseTemplate/releases/download/latest/echoverse_points.csv.gz")
```

 
```{r}
gg_bats <- ggplot(d, aes(x = x, y = y, color=bsize, image=image)) + 
  geom_image(aes(size=I(bsize)), alpha=1) +
  scale_color_gradient(low = "#194f68", high = "#56ffff") +   
  coord_cartesian(clip = "off") +
  labs(title = "echoverse") +  
  theme_void() +
  theme(plot.title = element_text(color = "#56ffff", size = 25,
                                  hjust = .5, vjust = 5, family = "Aller_Rg"), 
        legend.position = "none") 
 
# print(gg_bats)
```

## hexSticker 

```{r}    
s_size = 1
stick <- hexSticker::sticker(
  subplot = gg_bats, 
  #### Package name ####
  package = pkg, p_size=14, p_y = 1.4,
  #### Subplot #####
  s_x=1, s_y=.8,  s_height = s_size, s_width = s_size,
  #### Fill & border ####
  h_fill = "#25355c", h_color = "#41c6c8", #56ffff
  #### Spotlight ####
  spotlight = TRUE, l_alpha = .3, l_width = 10,
  #### File output ####
  filename = filename, dpi = 300)
print(stick)
```


# Session Info 

<details> 

```{r Session Info}
utils::sessionInfo()
```

</details>  

